### 2025-08-15 data-visualization.qmd
20250815 - Okabe-Ito color palette

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale). 
There are a variety of resources for color schemes that are accessible to color-blind viewers:

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> [@Kilin2022; archived at <https://perma.cc/7VTA-Y8YS>]
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- @Nunez2018

>>>>

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by [ColorBrewer](https://colorbrewer2.org) (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
[ColorBrewer](https://colorbrewer2.org) allows users to select colorblind safe, print friendly, and photocopy safe palettes.
Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available from the [`viridis` palette](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale). 
There are a variety of resources for color schemes that are accessible to color-blind viewers:

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> [@Kilin2022; archived at <https://perma.cc/7VTA-Y8YS>]
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- [Okabe-Ito color palette](https://jfly.uni-koeln.de/color/) [@Okabe2008; archived at <https://perma.cc/VD2E-XZZU>]
- @Nunez2018

The Okabe-Ito color palette [@Okabe2008] can be specified with the following colors:

```{r}
palette_OkabeIto <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999")

palette_OkabeIto_black <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#000000")
```

The Okabe-Ito color palette [@Okabe2008] is depicted in Figures [-@fig-okabeItoPaletteGray] and [-@fig-okabeItoPaletteBlack].

```{r}
#| label: fig-okabeItoPaletteGray
#| fig-cap: "Okabe-Ito Color Palette With Gray."
#| fig-alt: "Okabe-Ito Color Palette With Gray."

graphics::pie(rep(1, 8), col = palette_OkabeIto)
```

```{r}
#| label: fig-okabeItoPaletteBlack
#| fig-cap: "Okabe-Ito Color Palette With Black."
#| fig-alt: "Okabe-Ito Color Palette With Black."

graphics::pie(rep(1, 8), col = palette_OkabeIto_black)
```
